\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Ehat}{\widehat{E}}
\newcommand{\red}{\mathrm{red}}
\newcommand{\ord}{\mathrm{ord}}
\newcommand{\Aut}{\mathrm{Aut}}
\DeclareMathOperator{\Uk}{U}

\title{\bf When a Control Does Not Control:\\
A Mechanism-Derived Refutation of a False Positive\\
in $p$-adic Cryptanalysis of the ECDLP}

\author{W.\ T.\ Trevena\thanks{Correspondence: william.todd.trevena@gmail.com.
Code, data, and a per-phase index: the \texttt{ecdsa-lift} repository
(release tagged with this paper).}}

\date{June 2026}

\begin{document}
\maketitle

\begin{abstract}
Pipelines that fan out many structural probes against a hard problem and
surface the most anomalous are increasingly common, and they are
false-positive engines: with enough encodings and transforms an apparent
$5$--$11\sigma$ effect is cheap. We give a worked example from a
systematic negative exploration of $p$-adic lifting attacks on the
elliptic curve discrete logarithm problem (ECDLP) for ordinary curves of
the kind underlying ECDSA. A Gowers $\Uk^2$ / Fourier analysis of the
Teichm\"uller lift error $\delta(k)$ produced a persistent
$5$--$11\sigma$ anomaly on the original sweep (and $\sim$$3$--$14\sigma$
on a clean controlled set) that a standard permutation control certified
as ``real, index-ordered structure.'' We show this was a \emph{control that
cannot separate the hypotheses it was meant to test}. Two
\emph{mechanism-derived} nulls---one per candidate explanation---localize
the entire signal to a single exact, public identity (the antisymmetry
$\delta(k)+\delta(n-k)=[n]\tau(G)$): the signal vanishes when the
reflection is removed, and---our key positive control---synthetic
sequences carrying \emph{only} that reflection reproduce the elevation
and its growth with $n$ (and a power-law Fourier decay of the same order,
$\beta\approx0.4$). We then give an inertness theorem, an explicit
instance of a principle due to Silverman and to Gadiyar--Padma, that
explains why every such anomaly on these curves must reduce to
artifact-or-identity: for $\gcd(n,p)=1$ the only secret-dependent part of
any section lift error is $[k]c$, governed by $k \bmod p^{e-1}$, which is
not determined by the public data $(E,G,Q)$ over the natural lift and for
which no efficient algorithm from $(E,G,Q)$ is known. The
single inequality $\gcd(n,p)=1$ is sharp: when it fails ($n=p$, anomalous
curves) the obstruction vanishes and the same machinery recovers $k$ with
$100\%$ success---Smart's attack. The contribution is the methodology a
results-free negative exploration must practice: a control is valid only
against a named alternative hypothesis, nulls must be derived from
mechanisms, and a structural reason for inertness beats an accumulation
of passed tests.
\end{abstract}

\section{Introduction}

A recurring temptation in cryptanalysis is to lift a hard problem into a
richer setting---a $p$-adic ring, a formal group, a higher-genus
Jacobian---hoping that structure invisible over $\Fp$ becomes visible and
exploitable. For the ECDLP this hope is well mapped: Silverman's ``four
faces of lifting'' \cite{silverman4faces,silvermanlift} catalogues the
local/global $\times$ torsion/non-torsion strategies and the obstruction
each meets, and the only $p$-adic local lift that ever broke an ECDLP is
Smart's attack on anomalous curves \cite{smart,satoharaki,semaev}, where
the group order equals the characteristic.

Automating such explorations sharpens an old hazard. A pipeline that
computes many encodings of a lift invariant and reports the most unusual
will, after enough nulls, return an ``anomaly,'' and the analyst must
decide whether it is signal. We conducted a long, deliberately exhaustive
negative exploration of the canonical-lift attack surface for ordinary
curves with large prime-order subgroups---the ECDSA regime---and hit
exactly this case: a higher-order Fourier probe returned a clear
$5$--$11\sigma$ effect on the original sweep, a textbook control endorsed
it, and only a more careful experiment overturned it. The overturning, and the reason the
anomaly \emph{had} to be spurious, are the substance of this paper.

\paragraph{Contributions.}
\begin{itemize}
\item A worked, quantitative example of a \emph{control that does not
control}: a permutation test that appears to certify structure but
provably cannot separate the competing explanations
(\S\ref{sec:falsepos}).
\item \emph{Mechanism-derived controls}: enumerate the candidate
generating mechanisms and build one null per mechanism, including a
\emph{positive} control in which a synthetic model of the surviving
mechanism reproduces the statistic quantitatively
(\S\ref{sec:controls}). This localizes the anomaly (a $\sim$$3$--$14\sigma$
elevation on the clean controlled set) to one exact public identity.
\item An inertness theorem (\S\ref{sec:theorem}) explaining why such
anomalies are forced: for $\gcd(n,p)=1$ the secret-dependent part of any
section lift error is $[k]c$, governed by $k \bmod p^{e-1}$, which the
public data $(E,G,Q)$ does not determine over the natural lift and which
no known efficient algorithm computes from it. We present it as the explicit
instance it is of \cite{silverman4faces,gadiyarpadma}, with proof and
verification.
\item A sharp-boundary demonstration (\S\ref{sec:boundary}): the same
construction on an anomalous curve ($\gcd(n,p)=p$) recovers the secret
with $100\%$ success. The theorem's one hypothesis is exactly the line
between inert and broken.
\end{itemize}

The theorem is not new mathematics; its role is to certify, after the
fact, that the methodological suspicion was structurally warranted.

\section{Background and notation}\label{sec:background}

Let $p\ge 5$ be prime and $E/\Fp$ an elliptic curve with $\#E(\Fp)=N$.
Fix $G\in E(\Fp)$ of order $n$ with $\gcd(n,p)=1$, and $Q=kG$; the ECDLP
is to recover $k\in\Z/n\Z$ from $(E,G,Q)$. For $e\ge 1$, reduction modulo
$p$ on a fixed Weierstrass lift gives the exact sequence of abelian
groups
\begin{equation}\label{eq:ses}
0 \longrightarrow \Ehat(p\Z/p^e) \longrightarrow E(\Z/p^e)
  \xrightarrow{\ \red\ } E(\Fp) \longrightarrow 0,
\end{equation}
with kernel of reduction $\Ehat(p\Z/p^e)$ a $p$-group of order $p^{e-1}$~\cite{silvermanaec}.
The formal parameter $z=-X/Y$ identifies $\Ehat$ with $(p\Z/p^e,+_F)$
($+_F$ the formal group law), and the formal logarithm
$\log\colon\Ehat\to(p\Z/p^e,+)$ is an isomorphism with
$\log([m]u)=m\log(u)$. For $E\colon y^2=x^3+b$ the invariant differential
is $\omega=(1+3b\,z^6+\cdots)\,dz$, so
$\log(z)=z+\tfrac{3b}{7}z^7+\cdots$; for $z\in p\Z$ and $e\le 4$ the
correction vanishes ($z^7\in p^7\Z$), hence $\log(z)\equiv z\pmod{p^e}$
and kernel arithmetic is additive in $z$ at our precision.

A \emph{section} is a map $\tau\colon E(\Fp)\to E(\Z/p^e)$ with
$\red\circ\tau=\mathrm{id}$; the Teichm\"uller and Hensel lifts are
sections, generally not homomorphisms. We use the $x$-Teichm\"uller /
Hensel section throughout; it satisfies $\tau(-P)=-\tau(P)$ (the two
Hensel roots in $y$ are negatives), a property we invoke once below. The
\emph{lift error} (a $1$-cochain valued in $\Ehat$ by \eqref{eq:ses}) is
\begin{equation}
\delta_\tau(k)\ :=\ [k]\,\tau(G)\ -\ \tau(kG)\ \in\ \Ehat(p\Z/p^e).
\end{equation}
We write $\delta(k)$ for the Teichm\"uller case and
$z(\delta(k))\in p\Z/p^e$ for its formal parameter; for an integer
$z\in\Z/p^e$, $\mathrm{digit}_j(z)$ is the coefficient of $p^j$ in the
base-$p$ expansion (digit $0$ is the units place, identically $0$ here
since $z\in p\Z$).

\paragraph{Prior art.} That a $p$-adic lift transports the ECDLP without
weakening it, that the canonical (Teichm\"uller) lift carries no
information, and that ``solving the DLP is equivalent to constructing a
non-canonical lift'' are due to Silverman \cite{silverman4faces,
silvermanlift} for elliptic curves and to Gadiyar--Maini--Padma and
Gadiyar--Padma \cite{gadiyarpadma} for $\Fp^\times$. Smart, Satoh--Araki
and Semaev \cite{smart,satoharaki,semaev} give the anomalous attack; the
canonical lift is Lubin--Serre--Tate \cite{lst}, its computational
instance Satoh's point counting \cite{satohcount}. Our theorem repackages
these in the lift-error language.

\section{The exploration, in brief}\label{sec:exploration}

The full study ran forty numbered phases against the canonical-lift /
formal-group attack surface for $j=0$ (CM by $\Z[\omega]$) and generic
ordinary curves; Appendix~\ref{app:phases} indexes them. Probes included
pseudorandomness of $\delta$ (DFT flatness, maximal Berlekamp--Massey
complexity, uniform $p$-adic valuation, no polynomial/Mahler/%
multiplicative fit); a homomorphic section from $H^2$ vanishing (which
exists but, being a homomorphism, annihilates the attack); an explicit
Serre--Tate canonical lift validated by $\delta_{\mathrm{can}}\equiv 0$;
HNP/lattice (LLL, BKZ) attacks on the cocycle; MOV embedding degree;
ordinary/supersingular bridges; genus-$2$ Jacobian covers; Weil pairings;
Iwasawa $\mu$; and a Frobenius/unit-root analysis. Every one was inert.
Two exact relations survived, both \emph{public} (computable from the
curve alone) and both giving no advantage: an automorphism-equivariance
of the leading $\delta$-digit, and the antisymmetry
\begin{equation}\label{eq:antisym}
\delta(k)+\delta(n-k)=[n]\,\tau(G),
\end{equation}
which holds for any section with $\tau(-P)=-\tau(P)$ (so that
$\tau(kG)+\tau(-kG)=O$), and which we confirmed holds \emph{exactly} in
the $z$-coordinate---e.g.\ for $p=97$ ($n=79$),
$z(\delta(k))+z(\delta(n-k))$ takes a single value across all $39$
reflection pairs $(k,n-k)$.

\section{A false positive: the Gowers anomaly}\label{sec:falsepos}

Encode digit $j$ of the lift error as a phase,
$f_j(k)=\exp\!\big(2\pi i\,\mathrm{digit}_j(z(\delta(k)))/p\big)$ for
$k=1,\dots,n-1$, and measure the Gowers $\Uk^2$ norm~\cite{taovu} against an
$S^1$-uniform random baseline (definitions and counts in
\S\ref{sec:stats}). Across many primes the $\Uk^2$ norm was elevated
$5$--$11\sigma$ on the original exploratory sweep (Phase 37, a mixed bag
of curves later found to include supersingular and anomalous cases); on
the clean controlled set of \S\ref{sec:controls} the elevation is
$\sim$$3$--$14\sigma$ ($z(\text{real vs rand})$ ranges $3.3$ to $13.6$).
$\Uk^3$ and higher sat at baseline. The leading Fourier
coefficient decayed as $|\hat f(\alpha)|\sim C n^{-\beta}$ with
$\beta<1/2$, on CM, $j{=}1728$, and non-CM curves alike. An internal
report concluded this was a genuine, previously unnoticed degree-$1$
structural property of the Teichm\"uller cocycle ``not explained by any
universal identity.''

The evidence offered was a permutation control: randomly permuting the
\emph{values} of $f_j$ destroyed the elevation, read as proof that the
structure lived in the index ordering, hence was ``real.'' This inference
is invalid, and seeing why is the point of the paper.

\section{Mechanism-derived controls}\label{sec:controls}

\paragraph{The principle.} A control is a null model, and a null tests
exactly one alternative: the one under which the statistic is recomputed.
The value-permutation control recomputes $\Uk^2$ under ``the multiset of
digit values, in random order,'' so it separates ``structure in the value
distribution'' from ``structure in the ordering''---and nothing else. But
at least three mechanisms predict ordered structure here, and the
permutation kills all three at once, so it cannot attribute the signal to
any one:
\begin{enumerate}
\item[(H1)] \emph{Prefix-sum artifact.} $\delta(k)=-\sum_{i<k}c(iG,G)$ is
a running sum; summation is a low-pass filter
($\hat s(\xi)=\hat d(\xi)/(1-e^{2\pi i\xi/n})$, a $1/\xi$ pole at
$\xi=0$), so \emph{any} increments with this multiset, once summed, carry
elevated low-frequency $\Uk^2$.
\item[(H2)] \emph{Antisymmetry shadow.} The exact reflection
\eqref{eq:antisym} couples Fourier mode $\alpha$ with $-\alpha$, a rigid
constraint that elevates $\Uk^2$.
\item[(H3)] \emph{Genuine new structure}, the claimed explanation.
\end{enumerate}
We build one null per mechanism, each derived from how that mechanism
generates ordering. All numbers below are on the clean ordinary set of
Table~\ref{tab:p43} (Phase 43), under the pre-registered protocol of
\S\ref{sec:stats}.

\paragraph{H1 (increment shuffle).} $s(k)=z(\delta(k))$ is the exact
cumulative sum of its first differences $d(k)=s(k+1)-s(k)$. Shuffle $d$,
re-sum in the new order, recompute $\Uk^2$; the increment multiset is
identical, only the order changes. The shuffled-increment prefix sums sit
\emph{at} the random baseline ($z(\text{incShuf vs rand})\in[-0.3,+0.5]$
on every curve), while the true ordering sits $+3.5$ to $+12.5$ above
them, growing with $n$ (Table~\ref{tab:p43}, cols.\ 4--5). H1 is refuted:
the digit nonlinearity destroys the walk's low-frequency mass, so the
signal is in the \emph{specific} order, not in summation.

\paragraph{H2 (half-sequence).} The reflection \eqref{eq:antisym} relates
each $k$ to $n-k$. On the first half $k=1,\dots,\lfloor(n-1)/2\rfloor$
every reflection partner lies outside the window, so H2 predicts no
internal constraint there while H3 predicts the signal persists.
Re-running the increment-shuffle control on the half collapses the
elevation to noise ($z(\text{half vs incShuf})\in[-1.3,+1.5]$, no growth
in $n$; Table~\ref{tab:p43}, col.\ 6).

\paragraph{H2, positive control (the decisive test).} Absence of signal
on the half is necessary but not sufficient. We therefore test H2
\emph{constructively}: generate synthetic sequences that impose
\emph{only} the reflection, $s(k)$ i.i.d.\ uniform on the first half and
$s(n-k)=C-s(k)$ with $C=z([n]\tau(G))$, carrying no other structure, and
measure their $\Uk^2$. If reflection alone is the mechanism, these must
reproduce the real statistic. They do, quantitatively: the synthetic
$\Uk^2$ is elevated $+3.8$ to $+12.5$ over the random baseline---tracking
the real elevation across the whole range of $n$---and the real sequences
are statistically indistinguishable from them
($z(\text{real vs syn})\in[-1.4,+1.5]$; Table~\ref{tab:p43}, cols.\
7--8). The match on the $\Uk^2$ elevation and its growth with $n$ is the
substance of the positive control, and it is exact (cols.\ 7--8). The
finer digit-dependent Fourier decay is only \emph{partially} reproduced:
fitting $|\hat f|\sim C n^{-\beta}$ across the seven curves, the real
sequences give $\beta_{\mathrm{real}}\approx\{0.34,0.45,0.36\}$ for
digits $1,2,3$, whereas the reflection-only synthetic draws give a nearly
flat $\beta_{\mathrm{syn}}\approx\{0.41,0.41,0.40\}$. Reflection alone
thus reproduces the overall power-law decay ($\beta\approx0.4$, the same
order as the real exponents) but not its digit-to-digit variation; we do
\emph{not} claim the spectra match in detail. The decisive evidence that
H2 is the mechanism is the $\Uk^2$ match, which is genuine.

\paragraph{Verdict.} The anomaly is the Fourier-domain shadow of the
exact, public identity \eqref{eq:antisym}: H2, not H1 or H3. It is real
but not new, and---being a public consequence of the curve's geometry
that yields only a $2$-to-$1$ search reduction---carries no cryptanalytic
advantage. The varying ``decay exponent'' and its apparent curve-family
dependence in the internal report are finite-size scatter around a single
reflection-induced spectrum; nothing family-specific survives.

\begin{table}[t]
\centering
\small
\begin{tabular}{rrrrrrrr}
\toprule
& & & \multicolumn{3}{c}{negative controls ($z$)} & \multicolumn{2}{c}{positive ($z$)}\\
\cmidrule(lr){4-6}\cmidrule(lr){7-8}
$p$ & $n$ & $L$ & real/inc & inc/rand & half/inc & syn/rand & real/syn\\
\midrule
 67 &  79 &  78 & $+3.5..6.7$  & $-0.3..0.2$ & $-0.6..1.5$ & $+3.8..4.1$  & $-0.4..1.2$\\
163 & 139 & 138 & $+3.7..7.4$  & $+0.0..0.5$ & $-0.4..1.0$ & $+5.4..5.4$  & $-0.9..1.5$\\
211 & 199 & 198 & $+5.1..7.8$  & $+0.1..0.3$ & $-0.8..0.5$ & $+5.8..5.9$  & $-0.3..0.8$\\
349 & 313 & 312 & $+5.4..7.1$  & $-0.0..0.1$ & $-1.3..0.2$ & $+7.5..7.6$  & $-1.4..-0.4$\\
433 & 397 & 396 & $+8.2..9.6$  & $-0.1..0.2$ & $-0.4..0.6$ & $+8.4..8.6$  & $-0.1..0.7$\\
577 & 613 & 612 & $+7.8..11.5$ & $+0.1..0.2$ & $-0.5..1.1$ & $+10.2..10.4$& $-1.2..0.8$\\
823 & 829 & 828 & $+10.5..12.5$& $+0.1..0.2$ & $-0.9..0.0$ & $+12.2..12.5$& $-0.9..0.7$\\
\bottomrule
\end{tabular}
\caption{Phase 43, clean ordinary $j{=}0$ curves $y^2=x^3+7$
($p\equiv 1\bmod 3$, $n=\#E(\Fp)$ prime, so $\gcd(n,p)=1$ and the curve
is neither supersingular nor anomalous); $L$ is the number of usable
$k\in[1,n-1]$ after the pre-registered degeneracy filter
(\S\ref{sec:stats}); here $L=n-1$ (no drops). Ranges are over digits
$j=1,2,3$. Negative controls: shuffling increments leaves $\Uk^2$ at the
random baseline (inc/rand) and the true ordering far above it (real/inc),
which vanishes on the half (half/inc). Positive control: synthetic
reflection-only sequences reproduce the elevation (syn/rand) and the real
sequences match them (real/syn). The entire effect is the antisymmetry
\eqref{eq:antisym}.}
\label{tab:p43}
\end{table}

\section{Statistical protocol}\label{sec:stats}

All quantities in \S\ref{sec:controls} follow one pre-registered protocol
(fixed before the clean re-run; seed $20260609$). Precision $e=4$; digits
$j\in\{1,2,3\}$. The $S^1$-random baseline draws $300$ sequences
$g(k)=e^{i\theta_k}$, $\theta_k$ uniform on $[0,2\pi)$, of the same length
$L$; ``$\sigma$'' and ``$z$'' are the standard deviation and standardized
deviation \emph{of the relevant null}: incShuf uses $300$ increment
shuffles, syn uses $300$ synthetic draws, rand uses the $300$ baselines.
The Gowers $\Uk^2$ is computed exactly by its FFT formula. \emph{Exclusion
rule (pre-specified):} a value $z(\delta(k))$ is dropped iff its
projective lift is degenerate---$Y$ a non-unit, or $z\not\equiv 0\bmod p$
(a kernel element must have $z\equiv 0$); these are arithmetic failures of
the pure-Python lift, flagged automatically, and on the clean ordinary set
none occurred ($L=n-1$ throughout). The clean set is the seven ordinary
$j{=}0$ curves $y^2=x^3+7$ with $p\equiv 1\bmod 3$ and $n=\#E(\Fp)$ prime
(hence $\gcd(n,p)=1$, neither supersingular nor anomalous); the smallest
has $n=79$, so $L\ge 78$ throughout. No further size-based filter is
applied---every such curve in the run is reported.
We make no multiplicity correction and therefore do \emph{not} claim
discovery from any single elevated $\Uk^2$; the paper's logic is the
opposite---every elevation is traced to a named mechanism and shown to be
identity-or-artifact, and the controls are designed to \emph{remove}
signal, not to find it.

\section{Why it had to be a false positive: an inertness theorem}
\label{sec:theorem}

For $\gcd(n,p)=1$ the secret-dependent part of any section lift error is
forced into a coordinate the attacker cannot compute, so any
``structure'' in $\delta$ that \emph{is} computable from $(E,G,Q)$ must
reduce to a public identity or an artifact. We make this precise.

\begin{lemma}[Canonical lift]\label{lem:can}
If $\gcd(n,p)=1$, the order-$n$ subgroup $\langle G\rangle$ lifts
uniquely to $\langle\widetilde G\rangle\subset E(\Z/p^e)$ of order $n$
with $\red(\widetilde G)=G$, and $jG\mapsto[j]\widetilde G$ is a
homomorphism (so $\delta_{\tau_{\mathrm{can}}}\equiv 0$). Explicitly
$\widetilde G=[m]\tau(G)$ for any section $\tau$, where
$m\equiv 1\ (\bmod\,n)$, $m\equiv 0\ (\bmod\,p^{e-1})$.
\end{lemma}

\begin{proof}
$E(\Z/p^e)[n]$ injects into $E(\Fp)[n]$ (an $n$-torsion element of
$\Ehat$ is killed by $\gcd(n,p^{e-1})=1$, hence trivial) and surjects
onto it (Ă©tale $n$-torsion lifts by Hensel), so reduction is an
isomorphism on $n$-torsion and $\widetilde G$ is the unique order-$n$
lift of $G$. Writing $\tau(G)=\widetilde G+c$ with $c\in\Ehat$,
$[m]\tau(G)=[m\bmod n]\widetilde G+[m\bmod p^{e-1}]c=\widetilde G$.
\end{proof}

\begin{theorem}[Inertness]\label{thm:main}
Let $\tau$ be any section and $c:=\tau(G)-\widetilde G\in\Ehat$. Then:
\begin{enumerate}
\item[(a)] $\delta_\tau(k)=[k]c-c_k$, where $c_k:=\tau(kG)-[k]\widetilde G
\in\Ehat$ depends only on the point $kG$.
\item[(b)] $[k]c$ depends on $k$ only through $k\bmod p^{e-1}$;
$\log([k]c)=k\log(c)$ has additive order a power of $p$ dividing
$p^{e-1}$.
\item[(c)] $\gcd(n,p^{e-1})=1$, so by CRT, for $k$ uniform on
$\Z/np^{e-1}\Z$ the residues $k\bmod n$ (fixed by $Q$) and $k\bmod
p^{e-1}$ (governing $[k]c$) are independent. The public data $(E,G,Q)$
fixes only $k\bmod n$; hence over this natural lift $[k]c$, a function of
$k\bmod p^{e-1}$, is not determined by the public data, and no efficient
algorithm computing it from $(E,G,Q)$ is known.
\item[(d)] $\delta_\tau$ is a well-defined function of the instance
(i.e.\ of $k\bmod n$) iff $c=0$ (i.e.\ $\tau$ canonical on
$\langle G\rangle$), in which case $\delta_\tau\equiv 0$. For $c\neq 0$,
$\delta_\tau(k+n)-\delta_\tau(k)=[n]c=[n]\tau(G)\neq O$, the antisymmetry
constant of \eqref{eq:antisym}.
\end{enumerate}
\end{theorem}

\begin{proof}
(a) $[k](\widetilde G+c)-([k]\widetilde G+c_k)=[k]c-c_k$. (b) $c\in\Ehat$
has order dividing $p^{e-1}$; apply $\log$, whose order as a map
$k\mapsto k\log(c)$ on $(p\Z/p^e,+)$ is $p^e/\gcd(\log(c),p^e)$, a power
of $p$ dividing $p^{e-1}$ since $\log(c)\in p\Z$. (c) Since
$\gcd(n,p^{e-1})=1$, CRT gives $\Z/np^{e-1}\Z\cong\Z/n\Z\times\Z/p^{e-1}\Z$,
so for $k$ uniform on the left the residues $k\bmod n$ and $k\bmod
p^{e-1}$ are independent; as $(E,G,Q)$ determines only $k\bmod n$, it
leaves $k\bmod p^{e-1}$ (and hence $[k]c$) information-theoretically
unconstrained over this lift. We make no claim of a reduction beyond
this. (d) $c_k$ is a function of $kG$, hence of
$k\bmod n$; so $\delta_\tau$ factors through $k\bmod n$ iff
$k\mapsto[k]c$ does, iff $[n]c=O$, iff (as $\gcd(n,p^{e-1})=1$) $c=O$. The
shift equals $[n]c=[n]\tau(G)$ since $[n]\widetilde G=O$.
\end{proof}

\begin{corollary}\label{cor:main}
For $\gcd(n,p)=1$ no section lift error yields an efficiently computable
function of $(E,G,Q)$ that depends on the secret: it is either
identically zero (canonical) or not a function of the instance at all
(non-canonical, ambiguity $[n]\tau(G)\neq O$). The $p$-adic canonical
lift moves the ECDLP without weakening it.
\end{corollary}

This is why the Gowers anomaly was forced to be artifact-or-identity. The
only secret-bearing object, $[k]c$, is not publicly computable; the
visible $\delta$, taken as a function of the integer $k$, is either zero
or periodic-with-shift $[n]\tau(G)$---and that shift is the reflection
\eqref{eq:antisym} whose Fourier signature the probe rediscovered.

\paragraph{Remark (on clause (c)).} We deliberately do \emph{not} claim
$[k]c$ is information-theoretically independent of $(E,G,Q)$: once a
canonical representative $k\in[1,n-1]$ is fixed, $k\bmod p^{e-1}$ and
hence $[k]c$ are determined. The claim is computational---no efficient
algorithm produces $[k]c$ from the public data short of solving the
ECDLP---which is what matters for security and what the experiments probe.

\section{The boundary is sharp}\label{sec:boundary}

Theorem~\ref{thm:main} hinges on $\gcd(n,p)=1$. Set $n=p$ (an anomalous
curve, $\#E(\Fp)=p$): $k\bmod n=k\bmod p$ now coincides with the residue
the formal group sees, and the obstruction $[n]c=[p]c$ \emph{vanishes}
because $c\in\Ehat(p\Z/p^2)$ has order dividing $p$. The lift error
becomes well-defined and the formal logarithm of $[p]\cdot(\text{lift})$
is a homomorphism $E(\Fp)\to\Fp$ recovering $k$---Smart's attack. We
verified both sides (Table~\ref{tab:boundary}): on every ordinary test
curve the obstruction $[n]\tau(G)$ is a nonzero kernel element of
\emph{maximal} order $p^{e-1}$ (valuation $1$), while on the anomalous
$y^2=x^3+4x+7$ over $\mathbb F_{53}$ it vanishes and Smart's attack
recovers $k$ for $12/12$ random secrets. The same code path that is inert
for $\gcd(n,p)=1$ is a total break the instant $n=p$.

\begin{table}[t]
\centering
\small
\begin{tabular}{lcccc}
\toprule
Regime & curves & $\gcd(n,p)$ & obstruction $[n]\tau(G)$ & ECDLP\\
\midrule
Small ordinary, $\gcd(n,p)=1$ & $p=31$--$163$, $j{=}0$ & $1$ &
  nonzero, $\ord=p^{e-1}$ & inert\\
Anomalous (Smart) & $y^2{=}x^3{+}4x{+}7/\mathbb F_{53}$ & $p$ &
  $O$ (vanishes) & broken, $12/12$\\
\bottomrule
\end{tabular}
\caption{Phases 41--42. The well-definedness obstruction of
Theorem~\ref{thm:main}(d) is nonzero of maximal order on ordinary curves
and vanishes on anomalous curves, where Smart's attack succeeds.
$\gcd(n,p)=1$ is the exact boundary. Theorem~\ref{thm:main} is
regime-agnostic---it holds for any $\gcd(n,p)=1$---so these small ordinary
curves (chosen for exact-arithmetic reach) suffice to verify it; the
relevance to large-prime ECDSA curves is through the theorem's generality,
not through these particular orders.}
\label{tab:boundary}
\end{table}

\section{Numerical verification}\label{sec:verification}

Theorem~\ref{thm:main} was checked on $y^2=x^3+7$ at $e=4$ for ten
small ordinary curves over $\Fp$, $p=31,\dots,163$, all with
$\gcd(n,p)=1$ (Phase 41). These test orders are deliberately small and
several are composite---$n\in\{21,31,79,67,79,111,129,133,86,139\}$, five
of them composite---chosen for exact-arithmetic reach rather than to model
the large-prime ECDSA regime; since the theorem is regime-agnostic (it
needs only $\gcd(n,p)=1$), they suffice to verify it. Adding one
large-prime-order ordinary curve would strengthen the demonstration but is
not required for correctness. All kernel arithmetic is done through the
additive $z$-coordinate to avoid projective degeneracy on deep kernel
points. On every prime: the
obstruction $C_n:=z([n]\tau(G))$ is nonzero in $p\Z$ of valuation $1$ and
additive order exactly $p^{e-1}$ (clauses (d),(b)); the antisymmetry
$z(\delta(k))+z(\delta(n-k))\equiv C_n$ holds for all $k$ (clause (d) /
\eqref{eq:antisym}); the integer-representative shift
$z(\delta(k+n))-z(\delta(k))\equiv C_n$ holds on all eight
arithmetically-reliable primes (a group identity; on two primes the deep
multiply degenerates in exact arithmetic, flagged by a self-consistency
probe); and $\gcd(n,p^{e-1})=1$ (clause (c)). The canonical case
$\delta_{\mathrm{can}}\equiv 0$ (Lemma~\ref{lem:can}) was confirmed in an
exact $p$-adic computation. Anomalous recovery is Phase 42
(\S\ref{sec:boundary}).

\section{Discussion}\label{sec:discussion}

\paragraph{For negative cryptanalysis.} The reusable lesson is narrow and
practical. (i) A control tests exactly one alternative---the one its null
recomputes; before trusting it, enumerate the mechanisms that could
generate the statistic and check whether the control separates them. The
permutation test here failed because three mechanisms collapsed under it.
(ii) Build one null per mechanism, derived from how that mechanism
generates the statistic: H1 from the prefix-sum structure (shuffle
increments, re-sum), H2 from the reflection's reach (restrict to a half),
and---decisively---a \emph{positive} control synthesizing the surviving
mechanism (reflection-only sequences) and showing it reproduces the
statistic. (iii) Beware cumulative statistics: any running sum
manufactures low-frequency autocorrelation, so ``looks autocorrelated''
is the null, not the finding. (iv) Where a structural theorem is
available it bounds what controls can find; Theorem~\ref{thm:main} says no
honest control turns $\delta$ into an oracle for $\gcd(n,p)=1$.

\paragraph{For automated and AI-assisted exploration.} Pipelines that fan
out many probes and surface the most anomalous are false-positive
engines: with enough encodings and transforms an apparent $5$--$11\sigma$
effect is cheap, and a plausible-sounding control can launder it into a
``finding.'' The safeguard is not more probes but mechanism-derived nulls,
a positive control for any surviving signal, and---where possible---a
closed-form reason the probe must be inert. The theorem's value here is
exactly that: it turns ``we tried forty things and they failed'' into
``here is the one-line reason they had to.''

\paragraph{Honest scope.} Theorem~\ref{thm:main} is not new mathematics.
It is an explicit instance, in the lift-error language, of the principle
of \cite{silverman4faces,silvermanlift,gadiyarpadma} that lifting
transports the ECDLP; the anomalous degeneration is Smart's attack
\cite{smart}. It concerns section lift errors only; pairing-based
functionals are a separate mechanism (inert here for the independent
reason of huge embedding degree). We claim novelty only for the
methodology---the mechanism-derived controls and the positive control
that localizes the anomaly---and for the explicit identification of the
obstruction with the antisymmetry constant.

\section{Conclusion}

A long negative exploration produced one seductive anomaly. A standard
control endorsed it; mechanism-derived controls, capped by a positive
control, killed it and attributed it entirely to an exact public
identity; and an inertness theorem explains why no such anomaly on
ordinary curves can be more than artifact-or-identity, with the boundary
falling exactly at the anomalous curves where the lift attack genuinely
works. The cryptanalytic verdict is the expected negative one. The
transferable result is a discipline: name your alternative hypotheses,
build a null for each from its mechanism, synthesize the survivor to
confirm it, and prefer a structural reason for inertness over an
accumulation of passed tests.

\appendix
\section{Phase index}\label{app:phases}

\begin{table}[h]
\centering
\small
\begin{tabular}{rll}
\toprule
Phase(s) & Target & Verdict\\
\midrule
1--2   & injectivity / pseudorandomness of $\delta$ & injective, pseudorandom\\
3--4   & Smart--SSSA; homomorphic section from $H^2$ & section exists, attack does not\\
5--6   & CM-linearization; HNP on cocycle & pseudorandom; no bias\\
7--10  & MOV; poly fit; LLL coords; ord/ss bridge & inapplicable / no leak\\
11--13 & genus-2 Jacobian; GLV-CRT; Weil pairing & no new lever\\
14     & Aut$(E)$-equivariance rank law of $\delta$-digit & public, no advantage\\
15--20 & triple cocycle; DFT; Berlekamp--Massey; valuation; mult. & all inert\\
21b    & antisymmetry $\delta(k)+\delta(n-k)=[n]\tau(G)$ & exact, public\\
22--29 & coord change; formal log; Mazur--Tate; Mahler & no structure\\
30--36 & Sage validation; $2$-cocycle; canonical lift; Iwasawa & inert\\
37     & Gowers $\Uk^2$ anomaly & false positive (this paper)\\
40,40b & increment-shuffle and half-sequence controls & localizes to 21b\\
41     & numerical verification of Theorem~\ref{thm:main} & confirmed\\
42     & anomalous boundary (Smart) & $12/12$ recovery\\
43     & clean-ordinary re-run + positive control & confirms \S\ref{sec:controls}\\
\bottomrule
\end{tabular}
\caption{Index of the exploration. Full scripts, fixed seeds, and JSON
outputs are in the \texttt{ecdsa-lift} repository.}
\end{table}

\begin{thebibliography}{9}
\bibitem{silverman4faces} J.~H.~Silverman. \emph{The Four Faces of
Lifting for the Elliptic Curve Discrete Logarithm Problem.} Invited talk,
ECC 2007; slides available via the author's page. (Surveyed in
\cite{silvermanlift}.)
\bibitem{silvermanlift} J.~H.~Silverman. \emph{Lifting and Elliptic Curve
Discrete Logarithms.} Selected Areas in Cryptography (SAC) 2008, LNCS
5381, Springer, 2009, pp.\ 82--102.
\bibitem{gadiyarpadma} H.~G.~Gadiyar and R.~Padma. \emph{The discrete
logarithm problem over prime fields: the safe prime case. The Smart
attack, non-canonical lifts and logarithmic derivatives.} Czechoslovak Math.\ J.\ 68 (2018),
1115--1124 (arXiv:1702.07107); cf.\ Gadiyar, Maini, Padma, INDOCRYPT 2004,
LNCS 3348, 305--314.
\bibitem{smart} N.~P.~Smart. \emph{The discrete logarithm problem on
elliptic curves of trace one.} J.~Cryptology 12 (1999), 193--196.
\bibitem{satoharaki} T.~Satoh and K.~Araki. \emph{Fermat quotients and
the polynomial time discrete log algorithm for anomalous elliptic
curves.} Comment.\ Math.\ Univ.\ St.\ Pauli 47 (1998), 81--92.
\bibitem{semaev} I.~A.~Semaev. \emph{Evaluation of discrete logarithms on
some elliptic curves.} Math.\ Comp.\ 67 (1998), 353--356.
\bibitem{lst} J.~Lubin, J.-P.~Serre, J.~Tate. \emph{Elliptic curves and
formal groups.} Woods Hole Summer Institute, 1964 (lecture notes).
\bibitem{satohcount} T.~Satoh. \emph{The canonical lift of an ordinary
elliptic curve over a finite field and its point counting.} J.~Ramanujan
Math.\ Soc.\ 15 (2000), 247--270.
\bibitem{silvermanaec} J.~H.~Silverman. \emph{The Arithmetic of Elliptic
Curves}, 2nd ed., GTM 106, Springer, 2009.
\bibitem{taovu} T.~Tao and V.~Vu. \emph{Additive Combinatorics.}
Cambridge Univ.\ Press, 2006. (For the $\Uk^2$/Fourier correspondence.)
\end{thebibliography}

\end{document}
